Finite Time Analysis of Linear Stochastic Approximation

### 2017-04-30 flsa.tex
changes

We use $A\succeq 0$ to denote that the
square matrix $A$ is hermitian and positive semidefinite (HPD):
$A = A^\dag$, $\inf_x x^\dag A x\ge 0$.
For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
We also use $\preceq$ and $\prec$ analogously. $\E$ denotes mathematical expectation.

\begin{definition}
Let $P$ be a distribution over $\C_B^{\ncn}$, then define
\begin{align*}
A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
\end{align*}
Note that $\norm{A_P}<B$ and $\norm{C_P}<B^2$.
\end{definition}
\begin{definition}\label{contract}[Contraction Factors]
Define
\begin{align*}
\rhod{P}&\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}
\end{align*}
\end{definition}
Note that $\rhod{P}>\rhos{P}$.

\begin{definition}
We define a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$. We define a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} if all eigenvalues of $A$ have positive real parts.
\end{definition}
\begin{definition}\label{distpd}
Define a distribution $P\in \P$ to be \emph{positive definite} (PD)  if $A_P$ is positive definite. Define a distribution $P\in \P$ to be \emph{asymptotically stable} (AS) if $A_P$ is asymptotically stable.
\end{definition}

\section{Problem Setup}
We consider the linear stochastic approximation scheme with constant step size and PR-averaging given as below:
\begin{align}
\label{conststep} \theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
\label{iteravg} \thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_s,
\end{align}
where $\thh_t\in \C^{n}$ in \eqref{iteravg} is PR-average of the iterates $\theta_t\in \C^{n}$ of the constant step size stochastic recursion in \eqref{conststep}. For the quantities in \eqref{conststep,iteravg} we assume the following:
\begin{assumption}\label{lsadist}
$\{A_t,\, t\geq 0\}$ is an $i.i.d$ sequence of $\ncn$ matrices distributed according to $P\in \P$ and $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
\end{assumption}
From \cref{lsadist} it follows that $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ are martingale difference sequences with respect to the $\sigma$-field given by $\F_{t-1}=\{\theta_0,M_1,\ldots, M_{t-1}\}$.
Further, we assume that
\begin{assumption}\label{matvar}
%\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\end{assumption}

Our objective is to analyse $\EE{\norm{\thh_t-\theta^*}^2}$.

\begin{example}[Linear Prediction]
\end{example}

%\section{Error Recursion and Similarity Transformation}

\section{Main Results}
\begin{lemma}
Let $P$ be a positive definite distribution over $\C_B^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{theorem}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C_B^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

@@

\begin{theorem}[Change of Basis]
Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
\begin{align}
\EE{\norm{\gamma_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\begin{corollary}
Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}
\end{corollary}

\section{Uniform stepsizes}
\begin{definition}
Call a set of distributions $\P$ over $\C_B^{\ncn}$
\emph{weakly admissible} if there exists $\alpha>0$ such that
$\rhos{P}>0$ holds for all $P\in \P$.
\end{definition}
\begin{definition}
Call a set of distributions $\P$ over $\C_B^{\ncn}$ \emph{admissible}
if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
The value of $\alpha$ is called a witness.
\end{definition}
It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
hence if $\alpha>0$ witnesses that $\P$ is (weakly) admissible
then any $0<\alpha'\le \alpha$ is also witnessing this.

If $\P$ is weakly admissible, then one can choose a stepsize $\alpha>0$ solely based on the knowledge of $\P$ and
conclude that no matter the joint distribution of $(A_t,b_t)$, as long as the distribution of $A_t$ is an element of $\P$, the conclusions of \cref{maintheorem} hold. Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
A hopeful set is the following:
For $B>0$, let $\P_B$ is the set of distributions where for every $P\in \P_B$, $H_P+H_P^\top\succ 0$
and if $H\sim P$ then $\norm{H}\le B$ holds with probability one.
In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
Note that in most applications the knowledge of a suitable $B>0$ such that $\P\subset \P_B$ is available.
By appropriately normalizing the matrices involved, we may as well assume that $B=1$.

Our first result shows that we cannot be too ambitious:
\begin{proposition}
The set $\P_1$ is not weakly admissible.
\end{proposition}
\begin{proof}
Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$. Then $H_P = 2\epsilon I \succ 0$, hence $P\in \P_1$. Further, $Q_P = I$.
Hence, $\rho_\alpha(P) = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
\end{proof}
Note that the result still holds if in the definition of $\P_B$ we replace the condition that $H_P + H_P^\top \succ 0$ with the
stronger condition that $H_P\succ 0$. In words, positive definiteness of $H_P$ alone is not sufficient for weak admissibility.

In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
In particular, the random matrices $\{H_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
Define
\begin{align*}
\P_{\text{PSD},B} = \{ P \in \P_B \,:\,  \supp(P)\subset \text{PSD} \}\,.
\end{align*}
We have the following proposition:
\begin{proposition}
For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
and in particular any $0<\alpha < 2/B$ witnesses this.
\end{proposition}
\begin{proof}
Take any $P\in \P_{\text{PSD}}$ and let $H\sim P$.
Consider the SVD of $H$: $H = U \Lambda U^\top$ where $U$ is orthonormal and $\Lambda$ is diagonal with
nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
Then for any $x\in \R^n$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
Taking expectations we find that $x^\top Q_P x \le B x^\top H_P x$.
Hence, $\rho_\alpha(P) = 2 x^\top H_P x - \alpha x^\top Q_P x \ge (2- \alpha B ) \,x^\top H_P x $.
Thus, for any $\alpha<2/B$, $\rho_\alpha(P)>0$.
\end{proof}
\todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
I am guessing that for this they use that the loss is special and also that the noise is special.
This would be a good place to somehow incorporate this with a (short?) proof.
}

\begin{comment}
\paragraph{Contraction over time $t$}
The condition of contraction per time step is overly restrictive in the general case (other than the one presented in \Cref{rankonestep}). We can relax this condition and ask whether a weaker result in \Cref{multheorem} holds.
\begin{lemma}\label{multheorem}
Let $\P$ be the space of positive definite distributions supported on $\C_B^{\ncn}$. Then there exists an $\alpha_{\max}>0$ such that $\alpha\in(0,\alpha_{\max})$ implies $\us{\sup}{\norm{x}\leq 1,P\in\P,t\geq 0}\E_{P}[\norm{(I-\alpha H)^tx}]< \infty$.
\end{lemma}
\textbf{Possible Counter example for \Cref{multheorem} in $\R$}
Consider $\H=[-1,1]$ and $\P$ supported on $\{-1,1\}$ with $P(1)=\frac{1}{2}+\epsilon$ and $P(-1)=\frac{1}{2}-\epsilon$. Note that
\begin{align*}
\E[\Pi_{i=1}^t (1-\alpha H_i)]&=\exp{\ous{\sum}{i=1}{t}\log(1-\alpha H_i)}\\
&\approx \exp{t\E[\log(1-\alpha H_i)]}
\end{align*}
Now the condition $\E[\Pi_{i=1}^t (1-\alpha H_i)]<\infty,$ where $H_i\sim P$ loosely translates to the following
\begin{align*}
\E[\log(1-\alpha H_i)]<0,
\end{align*}
which can be verified by noting that
\begin{align*}
&\E[\log(1-\alpha H_i)]\\&=\E[(\frac{1}{2}+\epsilon)\log (1-\alpha)+(\frac{1}{2}-\epsilon)\log (1+\alpha)]\\
&=\frac{1}{2}\log(1-\alpha^2)+\epsilon(\log(1-\alpha)-\log(1+\alpha))\\
&<0
\end{align*}
\textbf{Possible Counter example for \Cref{multheorem} in $\R^2$}
Given any $\alpha$ the aim is to choose a distribution $P$ over $\H$ such that $\E_{P}[\norm{(I-\alpha H)^tx}]$ grows with $t$. In other words, given an $\alpha>0$, we are looking for a distribution $P'$ supported in the $\alpha$ neighbourhood of $I$ denoted by $\F$ such that $\E_{P'}[F^t]$ grows with $t$ (where $F\sim P'$). We now look at a possible suggestion for such a $P'$.\par
\begin{example}
Pick $F_1=\begin{bmatrix} 0 & 1-\beta_2 \\ 1+\beta_1& 0\\\end{bmatrix}^{\frac{1}{p}}$, $F_2=\begin{bmatrix} 0 & 1+\beta_1 \\ 1-\beta_2& 0\\\end{bmatrix}^{\frac{1}{p}}$, where $\beta_2>\beta_1$ such that $1+\frac{1}{2}(\beta_1-\beta_2)$. Let $P'$ be supported on $\{F_1,F_2\}$ with $P'(F_1)=\frac{1}{2}$ and $P'(F_2)=\frac{1}{2}$, and  $p$ is sufficiently large such that $F_1,F_2$ belong to the $\alpha$-neighbourhood of $I$. A possible way the random product can blow up is when products $F_2^pF_1^p$ occur in a contiguous stream. To see this,
\begin{align}
F_2^pF_1^p=\begin{bmatrix} (1+\beta_1) &0  \\ 0& (1-\beta_2)^2\\\end{bmatrix}
\end{align}
Thus $\norm{F_2^p F_1^p}>1$. However, such (i.e., $F_2^p F_1^p$) sequences have probability of occurance $\frac{1}{2^{2p}}$ is a sequence of length $2p$. Thus probability of them occuring quite often in a given random product is small.
\end{example}
\todoc[inline]{Did not we also want to discuss the multiplicative ergodic theorem? I have not cleaned this part yet..}
\end{comment}

>>>>

We use $A\succeq 0$ to denote that the
square matrix $A$ is hermitian and positive semidefinite (HPD):
$A = A^\dag$, $\inf_x x^\dag A x\ge 0$.
For $A,B$ HPD matrices, $A\succeq B$ holds if $A-B\succeq 0$.
We also use $A\succ B$ similarly to denote that $A-B \succ 0$.
We also use $\preceq$ and $\prec$ analogously. For any $x\in\C^n$x, we denote the general quadratic norm with respect to a $C\succ 0$ by $\norm{x}^2_C\eqdef x^\dag\, C \,x$.
$\E$ denotes mathematical expectation.

\begin{definition}
Let $P$ be a distribution over $\C^{\ncn}$, then define
\begin{align*}
A_P=\int M dP(M),\quad C_P=\int M^\dag M dP(M),
\end{align*}
\end{definition}
\begin{definition}\label{contract}[Contraction Factors]
Define
\begin{align*}
\rhod{P}&\eqdef {\inf}_{x\in\C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha A_P^\dag A_P\right)x},\\ \rhos{P}&\eqdef{\inf}_{x\in \C^n\colon\norm{x}=1}\ip{x,\left((A_P+A_P^\dag)-\alpha C_P\right)x}
\end{align*}
\end{definition}
Note that $\rhod{P}>\rhos{P}$.

\begin{definition}
We call a matrix $A\in \C^{\ncn}$ to be \emph{positive definite} (PD) if $\ip{x,Ax} >0,\,\forall x\neq 0 \in \C^{n}$. We call a matix $A\in \C^{\ncn}$ to be \emph{asymptotically stable} (AS) if all eigenvalues of $A$ have positive real parts. We call a matrix $A\in \C^{\ncn}$ to be \emph{symmetric positive definite} (SPD) is it is symmetric i.e., $A^\top=A$ and PD.
\end{definition}
\begin{definition}\label{distpd}
We call a distribution $P$ to be PD/AS/SPD if $A_P$ is PD/AS/SPD.
\end{definition}

\section{Problem Setup}
We consider the linear stochastic approximation (LSA) algorithm with constant step size and PR-averaging of the iterates.
\begin{subequations}\label{eq:lsa}
\begin{align}
\label{conststep}&\text{LSA:} &\theta_t&=\theta_{t-1}+\alpha(b_t-A_t\theta_{t-1}),\\
\label{iteravg}&\text{PR-Average:} &\thh_t&=\frac{1}{t+1}{\sum}_{i=0}^{t}\theta_i,
\end{align}
\end{subequations}
where $\thh_t, \theta_t \in \C^{n}$. The algorithm updates a pair of parameters $\theta_t,\tb_t\in \C^{\ncn}$ incrementally, in discrete time steps $t=1,2,\dots$
based on data $b_t\in \C^n$, $A_t\in \C^{n\times n}$\footnote{In almost all practical instances, the iterates $\theta_t,\thh_t\in \R^n$ and $A_t\in \R^{\ncn}, b_t\in \R^n$. However, it serves well for our analysis to make the general assumption that these qunatites are complex valued.}. Here $\alpha>0$ is a positive step-size parameter; the only tuning parameter of the algorithm besides the
initial value $\theta_0$. The iterate $\theta_t$ is treated as an internal state of the algorithm, while $\th_t$ is the output at time step $t$. The update of $\theta_t$ alone is considered a form of constant stepsize LSA. Note that the computation of $\theta_t$ can also be performed in an incremental fashion using $O(n)$ storage.

Owning to the special form of $A_t$, sometimes the matrix-vector product $A_t \theta_{t-1}$
can be efficiently computed in $O(n)$ time.
This happens for example when $A_t$ is rank one; examples of this kind arising in reinforcemnet learning \cite{} and linear prediction \cite{} will be presented in a later sections (\Cref{sec:rl,sec:lp}). The significance of efficient computation of the matrix-vector product is that then an update of the algorithm
can be implemented in $O(n)$ time and $O(n)$ storage, which makes the algorithm particularly attractive in large-scale computations when $n$ is in the range of thousands or more.
We are interested in the behaviour of \eqref{eq:lsa} under the assumptions presented in \Cref{assmp:lsa} In what follows, we make use of the $\sigma$-field given by $\F_{t-1}=\{\theta_0,A_1,\ldots, A_{t-1}, b_1,\ldots, b_{t-1}\}$.
\begin{assumption}\label{assmp:lsa}
\begin{enumerate}[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item \label{dist} $\{A_t\in\C^{\ncn},\, t\geq 0\}$ is an $i.i.d$ sequence distributed according to $P$ which is either AS or PD or SPD.
\item $\{b_t\}$ is an $i.i.d$ sequence of vectors in $\C^n$ with mean $\E[b_t]=b$.
\item \label{matvar} The martingale difference sequences\footnote{$\EE{M_t|\F_{t-1}}=0$ and $\EE{N_t|\F_{t-1}}=0$} $M_t\eqdef A_t-A_P$ and $N_t\eqdef b_t-b$ associated with $A_t$ and $b_t$ satisfy the following
%\begin{align*}\E\left[ M_t^\dag M_t\mid\F_{t-1}\right]\leq \Sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\begin{align*}\E\left[ \norm{M_t}^2\mid\F_{t-1}\right]\leq \sigma^2_P, \, \E[N_t^\dag N_t]=\sigma_b^2.\end{align*}
\item $A_P$ is invertible and there exists a $\ts=A_P^{-1}b$.
\end{enumerate}
\end{assumption}
\textbf{Performance Metric:}
The ``aim'' of the updates in \eqref{eq:lsa} is to estimate $\ts$ based on the noisy versions $(A_t,b_t)$ of $(A_P,b)$.
There are various ways of quantifying the error $\eh_t \doteq \thh_t- \ts$ of estimating $\ts$ using $\thh_t$, but the magnitude
of the errors is usually measured in some weighted quadratic norm. The expected loss for time step $t$ then becomes
\begin{align*}
L_t \doteq \EE{ \norm{\eh_t}_{C}^2 }\,,
\end{align*}
where recall that for a SPD matrix $C$, $\norm{x}_C^2 = x^\dag C x$. Here, the choice of the matrix $C$ depends on the application.\\
\textbf{Challenges:}
The idea of constant step size and iterate averaging is not new and goes back to \cite{}, where it was first shown that $L_t\ra 0$ as $t\ra\infty$. Recently \cite{bach} revisited the idea in connection to the stochastic gradient descent (SGD) algorithm (which is also an LSA of the form in \eqref{eq:lsa}) to solve linear regression. Much similar to the spirit of \cite{bach}, our aim here is to understand the aspects that effect the behaviour of $L_t$, such as:
\begin{itemize}
\item The variances $\sigma_P^2$ and $\sigma_b^2$ of the noise terms $M_t$ and $b_t$ in the update equation.
\item The baias $\norm{\theta_0-\ts}$ due to the initial condition $\theta_0$.
\item The effect of the structural properties of the $A_P$, i.e., whether it is AS or PD or SPD.
\end{itemize}
Most resutls in the past including those that of \cite{bach} consider an SGD algorithm for quadratic loss minimization, where the matrix $A_P$ is SPD and the noise has additional strucutral properties (we discuss this briefly in \Cref{sec:lp}). As it turns out these nice properties are lost when $A_P$ is AS or PD (a property that is holds in RL settings). While from the standpoint of the tools and techniques used, we use techniques much similar to \cite{bach}, however, as mentioned, the fact that we no longer are dealing with gradient systems (that involve to SPD matrices) but general vector fields (that involve AS or PD matrices), poses important challenges.
Many RL algorithms \cite{} are LSA algorithms that need only $O(n)$ updates. However, there is a lot of variability in their analyses. The analysis as well as the results in this paper are useful from the perspective that it presents a single unified way of understand the various LSA-RL algorithms. We now present and discuss the main results of this paper in the next section. The comparison with existing literature is done in two separate sections, one to compare our results with \cite{bach} in the linear predictions setting (\Cref{sec:lp}) and the other to compare our results with prior results in RL literature \cite{}.
\todoch{Mention impotance of forgetting the bias etc. Mention that we really don't have control over $\norm{U}$}
\section{Main Results}
\begin{lemma}
Let $P$ be a positive definite distribution over $\C^{\ncn}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{theorem}
Let $P$ in \Cref{lsadist} be a positive definite distribution over $\C^{\ncn}$. Then
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\theta^*}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}

@@

\begin{theorem}[Change of Basis]
Let $P$ be as in \Cref{lsadist}. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\theta^*$, and suppose if $P_U$ is positive definite, then
\begin{align}
\EE{\norm{\gh_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\begin{corollary}
Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
\begin{align}
\EE{\norm{\thh_t-\theta^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U}^2\norm{U^{-1}}^2}{\alpha \rhos{P_U}} \left(\frac{\norm{\theta_0-\theta^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\theta^*})}{t+1} \right)\,.
\end{align}
\end{corollary}

\section{Uniform stepsizes}
\begin{definition}
Call a set of distributions $\P$ over $\C^{\ncn}$
\emph{weakly admissible} if there exists $\alpha>0$ such that
$\rhos{P}>0$ holds for all $P\in \P$.
\end{definition}
\begin{definition}
Call a set of distributions $\P$ over $\C^{\ncn}$ \emph{admissible}
if there exists some $\alpha>0$ such that $\inf_{P\in \P} \rhos{P}>0$.
The value of $\alpha$ is called a witness.
\end{definition}
It is easy to see that $\alpha \mapsto \rhos{P}$ is decreasing,
hence if $\alpha>0$ witnesses that $\P$ is (weakly) admissible
then any $0<\alpha'\le \alpha$ is also witnessing this.

If $\P$ is weakly admissible, then one can choose a stepsize $\alpha>0$ solely based on the knowledge of $\P$ and
conclude that no matter the joint distribution of $(A_t,b_t)$, as long as the distribution of $A_t$ is an element of $\P$, the conclusions of \cref{maintheorem} hold. Further, if $\P$ is admissible then the error bound stated in  \cref{maintheorem} becomes independent of the instance.
Hence, an interesting question to investigate is whether a given set $\P$ is (weakly) admissible.
A hopeful set is the following:
For $B>0$, let $\P_B$ be the set of distributions over $\C_B^{\ncn}$ where for every $P\in \P_B$ is positive definite.
In what follows all distributions will be a subset of $\P_B$ with some $B>0$,
Note that in most applications the knowledge of a suitable $B>0$ such that $\P\subset \P_B$ is available.
By appropriately normalizing the matrices involved, we may as well assume that $B=1$.

Our first result shows that we cannot be too ambitious:
\begin{proposition}\label{notwad}
The set $\P_1$ is not weakly admissible.
\end{proposition}
\begin{proof}
Fix an arbitrary $\alpha>0$. We show that there exists $P\in \P$ such that $\rho_\alpha(P)<0$.
For $\epsilon \in (0,1/2)$ let $P$ be the distribution that is supported on $\{-I,I\}$ and takes on the value of $I$ with probability $1/2+\epsilon$. Then $A_P = 2\epsilon I \succ 0$, hence $P\in \P_1$. Further, $Q_P = I$.
Hence, $\rhos{P} = 4\epsilon-\alpha$. Hence, if $\epsilon<\alpha/4$, $\rho_\alpha(P)<0$.
\end{proof}
Thus \Cref{notwad} positive definiteness of $P\in \P_B$ alone is not sufficient for weak admissibility.
In linear least-squares problems (cf. \cref{ex:leastsquares}) the set $\P$ has more structure.
In particular, the random matrices $\{A_t\}$ are PSD. \todoc{The standard notation is $\mathrm{S}^{++}$ for the set of positive definite matrices. Introduce these and use them below..}
Define
\begin{align*}
\P_{\text{PSD},B} = \{ P \in \P_B \,:\,  \supp(P)\subset \text{PSD} \}\,.
\end{align*}
We have the following proposition:
\begin{proposition}
For any $B>0$, the set $\P_{\text{PSD},B}$ is weakly admissible
and in particular any $0<\alpha < 2/B$ witnesses this.
\end{proposition}
\begin{proof}
Take any $P\in \P_{\text{PSD}}$ and let $H\sim P$.
Consider the SVD of $H$: $H = U \Lambda U^\top$ where $U$ is orthonormal and $\Lambda$ is diagonal with
nonnegative elements. Note that $\Lambda \preceq B\, \I$ and thus $\Lambda^2 \preceq B \Lambda$.
Then for any $x\in \R^n$, $x^\top H^\top H x = x^\top U \Lambda^2 U^\top x \le B x^\top U \Lambda U^\top x = B x^\top H x$.
Taking expectations we find that $x^\top Q_P x \le B x^\top H_P x$.
Hence, $\rho_\alpha(P) = 2 x^\top H_P x - \alpha x^\top Q_P x \ge (2- \alpha B ) \,x^\top H_P x $.
Thus, for any $\alpha<2/B$, $\rho_\alpha(P)>0$.
\end{proof}
\todoc[inline]{So this altogether is weaker than what Bach founds. They prove strong admissibility.
I am guessing that for this they use that the loss is special and also that the noise is special.
This would be a good place to somehow incorporate this with a (short?) proof.
}
\subsection{Result Discussion}

For more details see the discussion on related work section

\section{Special Cases}
\subsection{Additive Noise Case}

\subsection{Symmetric Matrix}

\subsection{Linear Prediction With Structured Noise}

\section{Reinforcement Learning}

\section{Stochastic Approximation and Mirror Prox}

\begin{comment}
\paragraph{Contraction over time $t$}
The condition of contraction per time step is overly restrictive in the general case (other than the one presented in \Cref{rankonestep}). We can relax this condition and ask whether a weaker result in \Cref{multheorem} holds.
\begin{lemma}\label{multheorem}
Let $\P$ be the space of positive definite distributions supported on $\C^{\ncn}$. Then there exists an $\alpha_{\max}>0$ such that $\alpha\in(0,\alpha_{\max})$ implies $\us{\sup}{\norm{x}\leq 1,P\in\P,t\geq 0}\E_{P}[\norm{(I-\alpha H)^tx}]< \infty$.
\end{lemma}
\textbf{Possible Counter example for \Cref{multheorem} in $\R$}
Consider $\H=[-1,1]$ and $\P$ supported on $\{-1,1\}$ with $P(1)=\frac{1}{2}+\epsilon$ and $P(-1)=\frac{1}{2}-\epsilon$. Note that
\begin{align*}
\E[\Pi_{i=1}^t (1-\alpha H_i)]&=\exp{\ous{\sum}{i=1}{t}\log(1-\alpha H_i)}\\
&\approx \exp{t\E[\log(1-\alpha H_i)]}
\end{align*}
Now the condition $\E[\Pi_{i=1}^t (1-\alpha H_i)]<\infty,$ where $H_i\sim P$ loosely translates to the following
\begin{align*}
\E[\log(1-\alpha H_i)]<0,
\end{align*}
which can be verified by noting that
\begin{align*}
&\E[\log(1-\alpha H_i)]\\&=\E[(\frac{1}{2}+\epsilon)\log (1-\alpha)+(\frac{1}{2}-\epsilon)\log (1+\alpha)]\\
&=\frac{1}{2}\log(1-\alpha^2)+\epsilon(\log(1-\alpha)-\log(1+\alpha))\\
&<0
\end{align*}
\textbf{Possible Counter example for \Cref{multheorem} in $\R^2$}
Given any $\alpha$ the aim is to choose a distribution $P$ over $\H$ such that $\E_{P}[\norm{(I-\alpha H)^tx}]$ grows with $t$. In other words, given an $\alpha>0$, we are looking for a distribution $P'$ supported in the $\alpha$ neighbourhood of $I$ denoted by $\F$ such that $\E_{P'}[F^t]$ grows with $t$ (where $F\sim P'$). We now look at a possible suggestion for such a $P'$.\par
\begin{example}
Pick $F_1=\begin{bmatrix} 0 & 1-\beta_2 \\ 1+\beta_1& 0\\\end{bmatrix}^{\frac{1}{p}}$, $F_2=\begin{bmatrix} 0 & 1+\beta_1 \\ 1-\beta_2& 0\\\end{bmatrix}^{\frac{1}{p}}$, where $\beta_2>\beta_1$ such that $1+\frac{1}{2}(\beta_1-\beta_2)$. Let $P'$ be supported on $\{F_1,F_2\}$ with $P'(F_1)=\frac{1}{2}$ and $P'(F_2)=\frac{1}{2}$, and  $p$ is sufficiently large such that $F_1,F_2$ belong to the $\alpha$-neighbourhood of $I$. A possible way the random product can blow up is when products $F_2^pF_1^p$ occur in a contiguous stream. To see this,
\begin{align}
F_2^pF_1^p=\begin{bmatrix} (1+\beta_1) &0  \\ 0& (1-\beta_2)^2\\\end{bmatrix}
\end{align}
Thus $\norm{F_2^p F_1^p}>1$. However, such (i.e., $F_2^p F_1^p$) sequences have probability of occurance $\frac{1}{2^{2p}}$ is a sequence of length $2p$. Thus probability of them occuring quite often in a given random product is small.
\end{example}
\todoc[inline]{Did not we also want to discuss the multiplicative ergodic theorem? I have not cleaned this part yet..}
\end{comment}
